---
title: "Minderoo reporting data"
output: 
  github_document:
    html_preview: false
  word_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, echo = FALSE, dev =c("png","pdf"))
library(targets)
library(magrittr)
library(ggplot2)
library(ggtext)
library(ggdist)
library(ggchicklet)
library(cowplot)
library(gsheet)


period_dates <- lubridate::as_date(c("2020-06-01", "2021-06-13"))
reporting_dates <- lubridate::as_date(c("2017-06-01", "2021-06-13"))

label_date <- function(x){
  format(x, "%b") %>%
    stringr::str_replace("Jan", paste0("Jan\n", format(x, "%Y")))
}
set_date_limits <- function(x){
  c(lubridate::floor_date(x[1], "quarter"), 
    lubridate::ceiling_date(x[2], "quarter"))
}
```

```{r,,message = FALSE}
tar_load(landings)
tar_load(pds_trips)
tar_load(sites_url)
tar_load(species_url)

sites_codes <- gsheet::gsheet2tbl(sites_url) %>% 
  dplyr::rename(landing_site_name=site_code) %>% 
  dplyr::mutate(landing_site_name=as.character(landing_site_name))
species_codes <- gsheet::gsheet2tbl(species_url) %>% 
  dplyr::select(species_code,category) %>% 
  dplyr::rename(species=species_code) %>% 
  dplyr::mutate(species=as.character(species))

landings <-
  landings %>%
  dplyr::right_join(sites_codes) %>%
  dplyr::filter(!landing_site_name %in% c(35:39)) %>%# drop Conservation International sites
  dplyr::mutate(landing_site_name=site_name) %>%
  dplyr::select(-site_name)


landings <- landings %>%
  dplyr::mutate(date = lubridate::as_date(date), 
                in_period = date >= period_dates[1] & date <= period_dates[2]) %>%
  dplyr::filter(date >= reporting_dates[1] & date <= reporting_dates[2]) %>%
  dplyr::mutate(month = lubridate::floor_date(date, unit = "month")) %>%
  dplyr::mutate(dplyr::across(dplyr::ends_with("_fishers"), as.numeric)) %>%
  dplyr::mutate(fishers_total = 
                  `trip_group/no_fishers/no_child_fishers` +
                  `trip_group/no_fishers/no_men_fishers` + 
                  `trip_group/no_fishers/no_women_fishers`)

period_landings <- landings %>%
  dplyr::filter(in_period)

pds_trips <- pds_trips %>%
  dplyr::mutate(date = lubridate::as_date(Ended), 
                last_seen = lubridate::as_datetime(`Last Seen`, format = "%a %b %d %X UTC %Y", tz = "UTC"),
                in_period = date >= period_dates[1] & date <= period_dates[2], 
                `Duration (Seconds)` = as.numeric(`Duration (Seconds)`)) %>%
  # dplyr::filter(date >= reporting_dates[1], date <= reporting_dates[2]) %>%
  dplyr::mutate(month = lubridate::floor_date(date, unit = "month"), 
                month_last_seen = lubridate::floor_date(lubridate::as_date(last_seen), unit = "month")) 

period_pds <- pds_trips %>%
  dplyr::filter(in_period)

report_sites <- landings %>%
  dplyr::filter(in_period==TRUE) %>%
  magrittr::extract2("landing_site_name") %>% unique()

all_sites <- landings %>%
  dplyr::filter(in_period==FALSE) %>%
  magrittr::extract2("landing_site_name") %>% unique()

new_sites <-
setdiff(report_sites,all_sites)

```


- This report uses data of up to `r max(landings$date)` for the landing surveys and `r max(lubridate::as_date(pds_trips$Ended))` for the vessel tracking (Pelagic Data Systems).

- The reporting period in focus here is between `r period_dates[1]` and `r period_dates[2]`.

- The Minderoo funding helped us to continue and increase the amount of data collected and hence the quality of our estimates. Enumerators have recorded `r scales::number(nrow(landings))` landings so far, `r scales::number(nrow(period_landings))` of those during the reporting period. 

- The Minderoo funding also helped us to increase the geographic spread of the data collection. We were able to survey up to `r dplyr::n_distinct(period_landings$landing_site_name)` landing sites across Timor-Leste during the reporting period. 

- The increase in surveys and geographical spread allowed us to capture the work of `r scales::number(sum(period_landings$fishers_total, na.rm = T))` (non-unique) small-scale fishers during the reporting period. 

- Using the vessel trackers, we have tracked a total of `r dplyr::n_distinct(pds_trips$IMEI)` boats. The number of tracked boats has decreased over the reporting period. Presumably mainly due to deterioration, malfunction or physical damage to the devices.

- The trackers have recorded a total of `r scales::number(nrow(pds_trips))` trips, `r scales::number(nrow(period_pds))` of them during the reporting period. A smaller number than in previous periods of similar length. 

- Despite the smaller number of trips the effort of each fishing trip has remained relatively stable. 

```{r fig.height=4, fig.width=8}
landings %>%
  dplyr::count(month, in_period) %>%
  ggplot(aes(x = month, y = n)) +
  geom_col(aes(fill = in_period)) +
  scale_fill_manual(values = c("grey70", "#1a4985")) + 
  scale_x_date(date_breaks = "3 month", minor_breaks = NULL, labels = label_date, limits = set_date_limits) +
  labs(x = "", y = "Number of landings", 
       title = "The number of landings <strong>surveys increased</strong> during the <span style='color:#1a4985;'>reporting period</span>", 
       caption = "Monthly landings recorded by enumerators in Timor Leste") +
    theme_minimal() + 
  theme(legend.position = "none", 
        plot.title = element_textbox_simple(margin = grid::unit(c(0,0,1,0), "lines"))) 
```

```{r ,fig.height=4, fig.width=8}
landings %>%
  dplyr::group_by(month, in_period) %>%
  dplyr::summarise(landing_sites = dplyr::n_distinct(landing_site_name), .groups = "drop") %>%
  ggplot(aes(x = month, y = landing_sites)) +
  geom_col(aes(fill = in_period)) +
  scale_fill_manual(values = c("grey70", "#1a4985")) + 
  scale_x_date(date_breaks = "3 month", minor_breaks = NULL, labels = label_date, limits = set_date_limits) +
  labs(x = "", y = "Number of sites", 
       title = "The landings <strong>surveys's coverage was higher</strong> during the <span style='color:#1a4985;'>reporting period</span>", 
       caption = "Monthly number of landings sites surveyed by enumerators in Timor-Leste") +
    theme_minimal() + 
  theme(legend.position = "none", 
        plot.title = element_textbox_simple(margin = grid::unit(c(0,0,1,0), "lines"))) 
```

```{r,fig.height=4.5, fig.width=8}
landings %>%
  dplyr::group_by(month, in_period) %>%
  dplyr::summarise(dplyr::across(dplyr::ends_with("_fishers"), sum, na.rm = T), 
                   .groups = "drop") %>%
  tidyr::pivot_longer(dplyr::ends_with("_fishers")) %>%
  dplyr::mutate(name = stringr::str_replace(string = name, 
                                            pattern = stringr::regex('.*?([a-z]*)_[a-z]*$'),
                                            replacement = "\\1")) %>%
  dplyr::mutate(name = stringr::str_replace(string = name, 
                                            pattern = "child", 
                                            replacement = "children")) %>%
  dplyr::mutate(name = forcats::fct_reorder(name, value)) %>%
  ggplot(aes(x = month, y = value)) + 
  geom_col(aes(alpha = name, fill = in_period)) +
  guides(fill = "none", alpha = guide_legend(override.aes = list(fill = "#1a4985"))) +
  scale_fill_manual(values = c("grey70", "#1a4985")) +
  scale_alpha_manual(values = rev(c(1,0.8, 0.6))) +
  scale_x_date(date_breaks = "3 month", minor_breaks = NULL, labels = label_date, limits = set_date_limits) +
  theme_minimal()  + 
  labs(x = "", y = "Number of fishers", 
       title = "The number of <strong>fishers reached out was higher</strong> during the <span style='color:#1a4985;'>reporting period</span>", 
       caption = "Monthly number of (non-unique) fishers' involved in the surveyed landings by enumerators in Timor-Leste") +
    theme_minimal() + 
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
        plot.title = element_textbox_simple(margin = grid::unit(c(0,0,1,0), "lines")))
```

```{r, fig.height=4.5, fig.width=8}
label_date_price <- function(x){
  format(x, "%b") %>%
    stringr::str_replace("Feb", paste0("Feb\n", format(x, "%Y")))
}

landings %>%
  dplyr::filter(!is.na(total_catch_value) & survey_version %in% "v2") %>%
  ggplot(aes(x = month, y = as.numeric(total_catch_value))) +
  stat_pointinterval(aes(colour = in_period), .width = c(0.66, 0.9)) +
  scale_fill_manual(values = c("grey70", "#1a4985"), aesthetics = c("fill", "colour")) +
  scale_x_date(date_breaks = "3 month", minor_breaks = NULL, labels = label_date_price, limits = set_date_limits) +
  labs(x = "", y = "Catch value",
       title = "The <strong>total catch value has been stable</strong> during the <span style='color:#1a4985;'>reporting period</span>",
       caption = "Distribution of total catches value in Timor leste. Lines show the 33-66% and 10-90% percentiles.") +
  theme_minimal() +
  theme(legend.position = "none",
        plot.title = element_textbox_simple(margin = grid::unit(c(0,0,1,0), "lines")))
```

```{r, fig.height=4.5, fig.width=8,message = FALSE}
gear_type_groups <-landings %>%
  dplyr::filter(!is.na(`trip_group/gear_type`)) %>%
  dplyr::group_by(month,`trip_group/gear_type`) %>%
  dplyr::tally(name = "n_group")

landings %>%
  dplyr::filter(!is.na(`trip_group/gear_type`)) %>%
  dplyr::group_by(month) %>%
  dplyr::tally() %>%
  dplyr::left_join(gear_type_groups) %>%
  dplyr::mutate(n_perc=(n_group/n)*100,
                gear_type=`trip_group/gear_type`,
                gear_type=dplyr::case_when(
                  gear_type=="GN"~"Gill net",
                  gear_type=="HL"~"Hand line",
                  gear_type=="LL"~"Long line",
                  gear_type=="SG"~"Spear gun",
                  gear_type=="CN"~"Cast net",
                  gear_type=="MC"~"Manual collection",
                  gear_type=="BS"~"Beach seine",
                  gear_type=="SN"~"Seine net",
                  gear_type=="TP"~"Trap")) %>%
  ggplot(aes(x = month, y = n_perc,fill=gear_type)) +
  geom_chicklet(width = 27.5,alpha=0.7,na.rm = TRUE) +
  annotate("segment", as.Date("2020-06-01"), xend = as.Date("2021-06-13"), y = -1, yend = -1, colour = "#1a4985",
           size=1, alpha=0.8,arrow=arrow(ends = "both",length = unit(0.075, "inches")))+
  annotate("segment", as.Date("2020-06-01"), xend = as.Date("2021-06-13"), y = 101, yend = 101, colour = "#1a4985",
           size=1, alpha=0.8,arrow=arrow(ends = "both",length = unit(0.075, "inches")))+
  scale_x_date(date_breaks = "3 month", minor_breaks = NULL, labels = label_date, limits = set_date_limits) +
  ggthemes::scale_fill_tableau("Miller Stone", name = NULL,direction=1) +
  theme_minimal()  +
  labs(x = "", y = "Gear type usage (%)",
       title = "Relative usage of <strong>gear type  was comparable</strong> during the <span style='color:#1a4985;'>reporting period</span>",
       caption = "Percentage of different gear types usage on the monthly scale in Timor leste") +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        plot.title = element_textbox_simple(margin = grid::unit(c(0,0,1,0), "lines")))
```

```{r, fig.height=4.5, fig.width=8,message = FALSE}
habitat_groups <-landings %>%
  dplyr::filter(!is.na(`trip_group/habitat_boat`)) %>%
  dplyr::group_by(month,`trip_group/habitat_boat`) %>%
  dplyr::tally(name = "n_group")

landings %>%
  dplyr::filter(!is.na(`trip_group/habitat_boat`)) %>%
  dplyr::group_by(month) %>%
  dplyr::tally() %>%
  dplyr::left_join(habitat_groups) %>%
  dplyr::mutate(n_perc=(n_group/n)*100,
                habitat=`trip_group/habitat_boat`,
                habitat=dplyr::case_when(
                  habitat=="1"~"Reef",
                  habitat=="2"~"FAD",
                  habitat=="3"~"Deep",
                  habitat=="4"~"Beach",
                  habitat=="5"~"Traditional FAD",
                  habitat=="6"~"Mangrove",
                  habitat=="7"~"Seagrass")) %>%
  dplyr::mutate(n_perc=(n_group/n)*100) %>%
  ggplot(aes(x = month, y = n_perc,fill=habitat)) +
  geom_chicklet(width = 27.5,alpha=0.7,na.rm = TRUE) +
  annotate("segment", as.Date("2020-06-01"), xend = as.Date("2021-06-13"), y = -1, yend = -1, colour = "#1a4985",
           size=1, alpha=0.8,arrow=arrow(ends = "both",length = unit(0.075, "inches")))+
  annotate("segment", as.Date("2020-06-01"), xend = as.Date("2021-06-13"), y = 101, yend = 101, colour = "#1a4985",
           size=1, alpha=0.8,arrow=arrow(ends = "both",length = unit(0.075, "inches")))+
  scale_x_date(date_breaks = "3 month", minor_breaks = NULL, labels = label_date, limits = set_date_limits) +
  ggthemes::scale_fill_tableau("Tableau 20", name = NULL) +
  theme_minimal()  +
  labs(x = "", y = "Exploited habitat (%)",
       title = "<strong>Deep habitat was slightly more exploited </strong>during the <span style='color:#1a4985;'>reporting period</span>",
       caption = "Percentage of exploited habitat by fishers in Timor leste on the monthly scale") +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        plot.title = element_textbox_simple(margin = grid::unit(c(0,0,1,0), "lines")))
```

```{r, fig.height=9, fig.width=7.5,message = FALSE,warning=FALSE}

tot_catches <-
  landings %>%
  tidyr::unnest(cols=species_group) %>%
  dplyr::filter(!is.na(species) & !is.na(landing_site_name) & !species==0) %>%
  nrow()

catches_per_sites <-
  landings %>%
  tidyr::unnest(cols=species_group) %>%
  dplyr::filter(!is.na(species) & !is.na(landing_site_name) & !species==0) %>%
  dplyr::count(landing_site_name) %>%
  dplyr::rename(tot_catch_site=n)

species_stats <-
  landings %>%
  tidyr::unnest(cols=species_group) %>%
  dplyr::filter(!is.na(species) & !is.na(landing_site_name) & !species==0) %>%
  dplyr::mutate(n=as.numeric(n)) %>%
  dplyr::count(species) %>%
  dplyr::mutate(freq_species=(n/tot_catches)*100) %>%
  dplyr::arrange(desc(freq_species)) %>%
  dplyr::mutate(cumspe=cumsum(freq_species)) %>%
  dplyr::select(species,cumspe)


species_per_site <-
  species_per_site <-
  landings %>%
  tidyr::unnest(cols=species_group) %>%
  dplyr::filter(!is.na(species) & !is.na(landing_site_name) & !species==0) %>%
  dplyr::group_by(landing_site_name,species) %>%
  dplyr::count() %>%
  dplyr::left_join(species_stats) %>%
  dplyr::left_join(catches_per_sites) %>%
  dplyr::left_join(species_codes) %>%
  dplyr::mutate(species=category) %>%
  dplyr::select(-category)%>%
  dplyr::mutate(species=dplyr::case_when(cumspe >75 ~ "Other",
                                         TRUE~species)) %>%
  dplyr::mutate(n_perc=(n/tot_catch_site)*100) %>%
  dplyr::group_by(landing_site_name,species) %>%
  dplyr::summarise(n_perc=sum(n_perc),n=sum(n))


pdat <- species_per_site
pdat$landing_site_name <- as.factor(pdat$landing_site_name)

lev_data <-
  pdat %>%
  dplyr::select(landing_site_name,n) %>%
  dplyr::group_by(landing_site_name) %>%
  dplyr::summarise(n=sum(n)) %>%
  dplyr::arrange(desc(n)) %>%
  dplyr::select(landing_site_name) %>%
  magrittr::extract2("landing_site_name")

pdat$landing_site_name <- factor(pdat$landing_site_name,
                                 levels = lev_data)

site_col <- ifelse(lev_data %in% new_sites, "#1a4985", "grey50")
site_face <- ifelse(lev_data %in% new_sites, "bold", "plain")


p2 <-
  pdat %>%
  ggplot(aes(x = landing_site_name,y=n_perc,fill=species)) +
  geom_chicklet(alpha=1,na.rm = TRUE) +
  ggthemes::scale_fill_tableau("Tableau 20", name = NULL,direction = 1) +
  theme_minimal()  +
  labs(x = "Landing site", y = "Species distribution (%)")+
  theme_minimal() +
  coord_flip()+
  scale_x_discrete(limits=rev)+
  theme(legend.position = "",
        legend.title = element_blank(),
        axis.text.y = element_text(color = rev(site_col),face=rev(site_face)),
        plot.title = element_textbox_simple(margin = grid::unit(c(0,0,1,0), "lines")))

p1 <-
  pdat %>%
  dplyr::select(landing_site_name,n) %>%
  dplyr::group_by(landing_site_name) %>%
  dplyr::summarise(n=sum(n)) %>%
  ggplot(aes(x = landing_site_name,y=n)) +
  geom_point(aes(x = reorder(landing_site_name,n),y=n),
             alpha=0.5,na.rm = TRUE,color=site_col,fill=site_col,size=2) +
  geom_segment(aes(y = 0,x = landing_site_name,yend = n,xend = landing_site_name),
               alpha=0.25,color=site_col)+
  theme_minimal()  +
  labs(x = "", y = "Number of landings") +
  theme_minimal() +
  coord_flip()+
  theme(legend.position = "right",
        legend.title = element_blank(),
        plot.title = element_textbox_simple(margin = grid::unit(c(0,0,1,0), "lines")),
        axis.title.y = element_blank(),
        axis.text.y = element_blank())

legend_p2 <- get_legend(p2 +
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom"))

title <- ggdraw() + draw_label("Cairavela receives by far most of the landings. The species distribution\n is quite heterogeneous among the sites")
caption <- ggdraw() + draw_label("Distribution of species catches and landings among the different landing sites.\nNew landings sites are indicated in blue.",size=10)

p_grid <-plot_grid(p2,
            p1+theme(plot.margin = unit(c(0,0,0,-0.25), "cm")),
            rel_widths=c(3,0.75),align="h")

plot_grid(title,p_grid, legend_p2,caption, ncol = 1, rel_heights = c(0.1,1, 0.25,0.1))
```


```{r, fig.height=4.5, fig.width=8}
first_last_seen <- pds_trips %>%
  dplyr::group_by(IMEI) %>%
  dplyr::summarise(first_seen = min(month), 
                   last_seen = max(month_last_seen)) %>%
  tidyr::pivot_longer(dplyr::ends_with("seen"), values_to = "month")

month_seq <- function(x){
  seq.Date(min(x), max(x), "month")
}

first_last_seen %>%
  dplyr::group_by(IMEI) %>%
  tidyr::complete(IMEI, month = month_seq(month)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(name = "present") %>%
  dplyr::bind_rows(first_last_seen) %>%
  dplyr::mutate(in_period = month >= period_dates[1] & month <= period_dates[2]) %>% 
  dplyr::filter(!(name == "last_seen" & month == max(month))) %>%
  dplyr::mutate(name = stringr::str_replace_all(name, c(present = "Total", 
                                                        first_seen = "In", 
                                                        last_seen = "Out"))) %>%
  dplyr::count(month, in_period, name) %>%
  ggplot(aes(x = month, y = n)) + 
  geom_col(aes(fill = in_period)) +
  facet_grid("name", scales = "free_y", space = "free_y") +
  scale_fill_manual(values = c("grey70", "#1a4985"), aesthetics = c("fill", "colour")) + 
  scale_y_continuous(breaks = seq(0, 1000, by = 100)) +
  scale_x_date(date_breaks = "3 month", minor_breaks = NULL, labels = label_date, limits = set_date_limits) +
  labs(x = "", y = "Number", 
       title = "The number of <strong>boats tracked has decreased</strong> during the <span style='color:#1a4985;'>reporting period</span>", 
       caption = "Monthly number of boats with functioning tracking devices in Timor-Leste") +
  theme_minimal() + 
  theme(legend.position = "none",
        panel.spacing = grid::unit(1, "lines"),
        plot.title = element_textbox_simple(margin = grid::unit(c(0,0,1,0), "lines"))) 
```

```{r,fig.height=4.5, fig.width=8}
pds_trips %>%
  dplyr::count(month, in_period) %>%
  ggplot(aes(x = month, y = n)) +
  geom_col(aes(fill = in_period)) +
  scale_fill_manual(values = c("grey70", "#1a4985")) + 
  scale_x_date(date_breaks = "3 month", minor_breaks = NULL, labels = label_date, limits = set_date_limits) +
  labs(x = "", y = "Number of tracked trips", 
       title = "The number of <strong>trips recorded was lower</strong> during the <span style='color:#1a4985;'>reporting period</span>", 
       caption = "Monthly trips tracked with PDS trackers in Timor-Leste") +
    theme_minimal() + 
  theme(legend.position = "none", 
        plot.title = element_textbox_simple(margin = grid::unit(c(0,0,1,0), "lines"))) 
```

```{r, fig.height=4.5, fig.width=8}
pds_trips %>%
  ggplot(aes(x = month, y = `Duration (Seconds)`/3600)) +
  stat_pointinterval(aes(colour = in_period), .width = c(0.66, 0.9)) +
  scale_fill_manual(values = c("grey70", "#1a4985"), aesthetics = c("fill", "colour")) + 
  scale_x_date(date_breaks = "3 month", minor_breaks = NULL, labels = label_date, limits = set_date_limits) +
  labs(x = "", y = "Effort (hours)", 
       title = "The tracked trips' <strong>effort has been stable</strong> during the <span style='color:#1a4985;'>reporting period</span>", 
       caption = "Distribution of effort (hours) of tracked trips Timor-Leste. Lines show the 33-66% and 10-90% percentiles.") +
  theme_minimal() + 
  theme(legend.position = "none", 
        plot.title = element_textbox_simple(margin = grid::unit(c(0,0,1,0), "lines"))) 

```


